### 2021-11-16 documents/sections/extensions.tex
more work on yukpa

\subsection{\yukpa \obj{j-}}
\label{sec:yukpa}
The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms; it mostly employs innovative ones \parencite{meira2006syntactic}.
However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm is found \pcref{tab:yukreg}.
The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
%
\input{floats/yukreg}%
%
These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their reflex of the detransitivizer \detrz.
The same prefix occurs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}, contrasting with the reflex of \rc{u(j)-}, \obj{j(ɨ)-}.

In intransitive verbs, this first-person zero marking is also attested, but only for \obj{to} \qu{to go} \exref{yuk-7}.
That verb contrasts with regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
Thus, we conclude that the extension of \obj{j-} in \yukpa stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.

>>>>

\subsection{\yukpa \obj{j-}}
\label{sec:yukpa}
The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm is found \pcref{tab:yukreg}.
The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
%
\input{floats/yukreg}%
%
These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their reflex of the detransitivizer \detrz.
The same prefix occurs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we identify the zero marking in \gl{1}>\gl{3} scenarios as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.

In intransitive verbs, this first-person zero marking is also attested, but only for \obj{to} \qu{to go} \exref{yuk-7}.
That verb contrasts with regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
In summary, the extension of \obj{j-} in \yukpa only spared \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
